---
title: "Cohen's u3"
format: html
server: shiny
---

## Overview

This application is designed to communicate effect size Cohen's $d$ for non-technical audiences based on [Alliger's (2024) explanation](https://drive.google.com/file/d/1--777v4s9FbWg8gyfZiXwNCfXnkI4MOw/view){target="_blank"} using Cohen's $U_3$ ([Cohen, 1988, p. 20-23](https://www.utstat.toronto.edu/~brunner/oldclass/378f16/readings/CohenPower.pdf){target="_blank"}). Input the pre-training success rate and training effect size $d$ below to produce the Cohen's $U_3$ value for the post-training improvement in performance.

```{r}

numericInput(inputId = "base_rate", 
             label = "Pre-Training Success Rate:",
             value = .5,
             min = 0,
             max = 1,
             step = .01)
numericInput(inputId = "d", 
             label = "Cohen's d:",
             value = .53,
             step = .01)
br()
checkboxInput(inputId = "detail", 
             label = "Detailed output",
             value = FALSE)
```

## Result
```{r}
textOutput("text")
verbatimTextOutput("console")
br()
plotOutput("plot")
```

```{r}
#| context: server
#| warning: false
#| message: false
library(ggplot2)
library(tidyverse)
library(shiny)

## Text explanation render

nround <- function(values, digits) {
  factor <- 10^digits
  is_neg <- ifelse(values < 0, -1, 1)
  result <- is_neg*floor(abs(values)*factor+.5)/factor
  return(result)
}

output$text <- renderText({

  z_pre  <- qnorm(input$base_rate)
  z_post <- z_pre + input$d
  pdf_post <- pnorm(z_post)
  pct_imp <- pdf_post - input$base_rate
  bsr_scl <- nround(input$base_rate*100,0)
  pct_scl <- nround(pct_imp*100,0)
  msg <- paste0("Relative to the pre-training success rate of ", bsr_scl ,"%, successful performance post-training increased ", pct_scl, "%")
  return(msg)

})

## Detailed output render

output$console <- renderText({
  # Values
  z_pre  <- qnorm(input$base_rate)
  z_post <- z_pre + input$d
  pdf_post <- pnorm(z_post)
  
  u2 <- pnorm(input$d/2)
  u1 <- (2*u2-1)/u2
  u3 <- pdf_post #Good
  u3_inc <- pdf_post - input$base_rate #Good
  r <- input$d/sqrt(input$d^2+4)
  
  # Formatting
  bsr_scl <- nround(input$base_rate*100,0)
  u1_scl <- nround(u1*100,1)
  u2_scl <- nround(u2*100,1)
  u3_scl <- nround(u3*100,1)
  u3_inc_scl <- nround(u3_inc*100,1)
  r_scl <- nround(r,3)
  R2_scl <-  nround(r^2,3)
  
  
  if (input$detail == TRUE)  {
    return(print(paste0(
      "\n",
      "INPUTS\n",
      "-------------------------------\n",
      "Pre-training success rate: ", bsr_scl,"%\n",
      "Cohen's d: ", input$d,"\n\n",
      "OUTPUTS\n",
      "-------------------------------\n",
      "Cohen's U1:  ", u1_scl,"%\n",
      "Cohen's U2:  ", u2_scl,"%\n",
      "Cohen's U3:  ", u3_scl,"%\n",
      "Increase U3: ", u3_inc_scl,"%\n",
      "Pearson's r: ", r_scl,"\n",
      "R-squared:   ", R2_scl,"\n"
      )))
  }
})

## Plot Render
output$plot <- renderPlot({

z_values <- seq(-3, 3, length.out = 100)
d_val <- .5

null_dist <- dnorm(z_values)
alt_dist <- dnorm(z_values - input$d)  # Cohen's d = 0.5 shift

data <- data.frame(
  z_values = rep(z_values, 2),
  density = c(null_dist-.01, alt_dist-.01),
  group = factor(rep(c("H0", "Ha"), each = length(z_values)))
)

## Data for the annotation layer

z_pre  <- qnorm(input$base_rate)
z_post <- z_pre + input$d
pdf_post <- pnorm(z_post)

data_ann <- data.frame(group = factor(c("H0", "Ha")),
                       labels = c(
                         paste0("Pre-training\nsuccess rate: ", nround(input$base_rate*100, digits = 0), "%"),
                         paste0("Post-training\nsuccess rate: ", nround(pdf_post*100, digits = 0), "%")
                       )
                       )

ggplot(data, aes(x = z_values, y = density, fill = group, color = group)) +
  geom_ribbon(data = data %>% filter(group == "Ha" & z_values >= qnorm(1-input$base_rate)), # Pass rate in post
              aes(ymin = 0, ymax = density)) +  # Power region
  geom_ribbon(data = data %>% filter(group == "H0" & z_values >= qnorm(1-input$base_rate)),
              aes(ymin = 0, ymax = density)) +  # Base Rate in Pre
  geom_line(data = data %>% filter(group == "H0"), size = 1) + 
  geom_line(data = data %>% filter(group == "Ha"), size = 1) + 
  geom_text(data = data_ann, x = -2, y = .25, aes(label = labels), size = 8) +
  scale_fill_manual(values = c("#336666","#FF7F7F")) +
  scale_color_manual(values = c("#336666","#FF7F7F")) +
  scale_x_continuous(expand = c(0,0), limits = c(-3,3), breaks = c(seq(-3,3, 1))) +
  scale_y_continuous(expand = c(0,0), limits = c(0,.4)) +
  facet_grid(rows = "group") +
  theme_void() +
  theme(
    strip.text = element_blank(),
    legend.position = "none"
  )

})

```

## Code

This is the code for the function above. Note, I added a custom function to round to the nearest integer since this is not the default behavior in R's round() function. It is not required for the u3 function.

```{r}
#| echo: true
nround <- function(values, digits) {
  factor <- 10^digits
  is_neg <- ifelse(values < 0, -1, 1)
  result <- is_neg*floor(abs(values)*factor+.5)/factor
  return(result)
}

u3 <- function(base_rate, d, interpret = TRUE) {
  z_pre  <- qnorm(base_rate)
  z_post <- z_pre + d
  pdf_post <- pnorm(z_post)
  pct_imp <- pdf_post - base_rate
  pct_scl <- nround(pct_imp*100,0)
  msg <- paste0("Relative to the pre-training group, the training resulted in a ", pct_scl, "% increase in successful performance.")
  if (interpret == TRUE) {
    return(msg)
  }
  if (interpret == FALSE)  {
    return(pct_imp)
  }
}
```

## References

Alliger, G. M. (2024). How to better communicate the likely effectiveness of proposed training: Interpreting meta-analytic results for (and with) an applied audience. *International Journal of Training Research*, *0*(0), 1–8. [https://doi.org/10.1080/14480220.2024.2397355](https://doi.org/10.1080/14480220.2024.2397355){target="_blank"}

Cohen, J. (1988). *Statistical power analysis for the behavioral sciences* (2nd ed.). Lawrence Erlbaum Associates, Publishers.

## Resources

- [R package for calculating Cohen's $U_3$ using raw data](https://easystats.github.io/effectsize/reference/p_superiority.html){target="_blank"}
- [Kristoffer Magnusson's description of Cohen's $d$ and $U_3$](https://rpsychologist.com/cohend/){target="_blank"}
